### 2008-12-17 talk/ecoop2009/benchmarks.tex
move the paragraph about tlc features to the TLC section, and add the benchmarks written in C#

\cfbolz{I think this should go to the beginning of the description of the TLC as
it explains why it is written as it is written}

Despite being very simple and minimalistic, \lstinline{TLC} is a good
candidate as a language to run benchmarks, as it has some of the features that
makes most of current dynamic languages so slow:

\begin{itemize}

\item \textbf{Stack based VM}: this kind of VM requires all the operands to be
  on top of the evaluation stack.  As a consequence programs spend a lot of
  time pushing and popping values to/from the stack, or doing other stack
  related operations.  However, thanks to its simplicity this is still the
  most common and preferred way to implement VMs.

\item \textbf{Boxed integers}: integer objects are internally represented as
  an instance of the \lstinline{IntObj} class, whose field \lstinline{value}
  contains the real value.  By having boxed integers, common arithmetic
  operations are made very slow, because each time we want to load/store their
  value we need to go through an extra level of indirection.  Moreover, in
  case of a complex expression, it is necessary to create many temporary
  objects to hold intermediate results.

\item \textbf{Dynamic lookup}: attributes and methods are looked up at
  runtime, because there is no way to know in advance if and where an object
  have that particular attribute or method.
\end{itemize}

In the following sections, we will show some benchmarks that show how our
generated JIT can handle all the features above very well.

@@

@@

As we can see, the code generated by the JIT is almost 500 times faster than
the non-jitted case, and it is only about 1.5 times slower than the same
algorithm written in C\#, which can be considered the optimal goal.

>>>>

In section \ref{sec:tlc-features}, we saw that TLC provides most of the
features that usaully make dynamically typed language so slow, such as
\emph{stack-based VM}, \emph{boxed arithmetic} and \emph{dynamic lookup} of
methods and attributes.

In the following sections, we will show some benchmarks that show how our
generated JIT can handle all these features very well.

@@

Moreover, for each benchmark we also show the time taken by running the
equivalent program written in C\#.  By comparing the results against C\#, we
can see how far we are from the supposed optimal performances.  \anto{I
  don't really like the last sentence, but right now I can't think of another
  way to phrase it.  Rewording welcome :-)}

@@

As we can see, the code generated by the JIT is almost 500 times faster than
the non-jitted case, and it is only about 1.5 times slower than the same
algorithm written in C\#: the difference in speed it is probably due to both
the fact that the current CLI backend emits slightly non-optimal code and that
the underyling .NET JIT compiler is highly optimized to handle bytecode
generated by C\# compilers.

### 2008-12-17 talk/ecoop2009/tlc.tex
move the paragraph about tlc features to the TLC section, and add the benchmarks written in C#

>>>>

\subsection{TLC features}
\label{sec:tlc-features}

Despite being very simple and minimalistic, \lstinline{TLC} is a good
candidate as a language to test our JIT generator, as it has some of the
features that makes most of current dynamic languages so slow:

\begin{itemize}

\item \textbf{Stack based VM}: this kind of VM requires all the operands to be
  on top of the evaluation stack.  As a consequence programs spend a lot of
  time pushing and popping values to/from the stack, or doing other stack
  related operations.  However, thanks to its simplicity this is still the
  most common and preferred way to implement VMs.

\item \textbf{Boxed integers}: integer objects are internally represented as
  an instance of the \lstinline{IntObj} class, whose field \lstinline{value}
  contains the real value.  By having boxed integers, common arithmetic
  operations are made very slow, because each time we want to load/store their
  value we need to go through an extra level of indirection.  Moreover, in
  case of a complex expression, it is necessary to create many temporary
  objects to hold intermediate results.

\item \textbf{Dynamic lookup}: attributes and methods are looked up at
  runtime, because there is no way to know in advance if and where an object
  have that particular attribute or method.
\end{itemize}
